---
title: ""
output: pdf_document
geometry: margin=0.5cm
params:
  reactive_df: NA
  chosen_sex: NA
header-includes:
- \usepackage[default]{sourcesanspro}
- \usepackage[T1]{fontenc}
- \usepackage{fancyhdr}
- \usepackage{graphicx}
- \usepackage{longtable}
- \usepackage{tabu}
- \usepackage{booktabs}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{hyperref}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
mainfont: SourceSansPro
---

```{r, include=FALSE}

options(tinytex.verbose = TRUE)

```

```{r echo=FALSE, message= FALSE}

library(ggplot2)
library(dplyr)
library(kableExtra)
library(tidyr)
library(knitr)
library(stringr)

mydate <- Sys.Date()

```



\blandscape

  \begin{minipage}{0.7\textwidth}
    \raggedright
    \textbf{\large{PHS Adult Mental Health Indicators}} \\
    \large{National profile for Scotland}\\
    Sex: `r params$chosen_sex` \\
    Downloaded on: `r mydate` \\
    From: Public Health Scotland's Adult Mental Health Indicators dashboard \\
  \end{minipage}
  \hfill
  \begin{minipage}{0.3\textwidth}
    \raggedleft
    \includegraphics[width=100pt]{phs-logo-updated.png}
  \end{minipage}
  \bigskip
  \begin{minipage}{0.95\textwidth}
    \raggedright
    \bigskip
    \large{This national profile shows the latest available mental health indicator data for the population of Scotland, and compares the most and least deprived areas.} \par 
    \bigskip
    \large{Please see the dashboard for more detailed information, including data sources and data downloads.}  
  \end{minipage}

```{r setup, include=FALSE, dev="cairo_pdf"} 


knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")


```


```{r echo=FALSE, message= FALSE}


# function to create sparkline for each indicator
create_sparkline_nat <- function(data, col1, col2, maxval) {
  plot <- ggplot(data, aes(x=1:length(trend), y=as.numeric(trend))) +
    geom_line(linewidth=2, colour=col1) +
    geom_area(fill=col2) +
    theme_void() +
    guides(fill = "none") +
    scale_y_continuous(limits=c(0, maxval), expand = expansion(mult = c(0, 0.2))) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
          plot.background = element_rect(fill = "white", color = "white"),
          panel.background = element_rect(fill = "white", color = "white"))
  
  return(plot)
}

plot_nodata <- function(height_plot = 20) {
  df<-data.frame(x=1,y=1)
  ggplot(df,aes(x,y)) +
    geom_blank() +
  #  annotate("text", x = 1, y = 1, size = 12, label = "No data") +
    theme_void() 
}

```


```{r echo=FALSE, message= FALSE}

# temporary folder to save charts to
graph_directory <- tempdir()


```


```{r echo=FALSE, message= FALSE}
# # the data to plot
df <- params$reactive_df %>%
  select(domain, indname = ind_name, sdef = short_definition, yrange=year_range, 
         value, mdvalue = most_depr_value, ldvalue = least_depr_value, maxval = max_val,
         trend, leastdeptrend, mostdeptrend) %>%
  mutate(value = str_replace_all(value, "%", "\\\\%"),
         value = str_replace_all(value, "\n", ""),
         mdvalue = str_replace_all(mdvalue, "%", "\\\\%"),
         mdvalue = str_replace_all(mdvalue, "\n", ""),
         ldvalue = str_replace_all(ldvalue, "%", "\\\\%"),
         ldvalue = str_replace_all(ldvalue, "\n", ""))

```


```{r echo=FALSE, message= FALSE}
# the data to be displayed in table
table <- df %>%
  mutate(indname = ifelse(indname=="Alcohol-related hospitalisation (mental and behavioural disorders)", "Alcohol-related hospitalisation*", indname)) %>%
  mutate(domain = gsub(" ", "\n", domain)) %>%
  mutate(domain = linebreak(domain)) %>%
  mutate(yrange = gsub("to", "to\n", yrange)) %>%
  mutate(yrange = linebreak(yrange, align = c("c"))) %>%
  mutate(value = gsub("/100,000", "\n/100,000", value)) %>%
  mutate(mdvalue = gsub("/100,000", "\n/100,000", mdvalue)) %>%
  mutate(ldvalue = gsub("/100,000", "\n/100,000", ldvalue)) %>%
  mutate(value = linebreak(value)) %>%
  mutate(mdvalue = linebreak(mdvalue)) %>%
  mutate(ldvalue = linebreak(ldvalue)) %>%
 # mutate(ind = paste0(indname, "\n", sdef, " ", yrange, ".")) %>%
 # mutate(ind = linebreak(ind, double_escape = TRUE)) %>% # "! Paragraph ended before \mcell@tabular was complete." / "! Missing } inserted."
 # mutate(ind = paste0(
 #   "\\textbf{",indname, "}\n", "\\textit{", sdef, " ", yrange, ".}")) %>%
 # mutate(ind = linebreak(ind)) %>% # "! Paragraph ended before \mcell@tabular was complete." / "! Missing } inserted."
  # mutate(indname2 = paste0(
  #   "\\textbf{",indname, "}\n", "\\textit{", sdef, ".}")) %>%
  # mutate(indname2 = paste0("\\makecell[c]{",
  #                          indname, "\\\\", 
  #                          sdef, " ", yrange, ".}")) %>%
  # mutate(indname3 = paste0(indname, "\\\\", 
  #                          sdef, " ", yrange, ".")) %>%
  mutate(id = row_number(),
         place2 = "A",
         place3 = "B") %>% #placeholders, so there are enough cols
  select(domain, indname, yrange, value, id, 
         mdvalue, place2, 
         ldvalue, place3)
#table$ind <- linebreak(table$ind) 
#\makecell[c]{a\\b\\c}
# when table$ind not commented out: Warning: Error in <Anonymous>: arguments imply differing number of rows: 8, 2, 6; or (error not seen recently)
# Paragraph ended before \mcell@tabular was complete. (ind and indname2 too)
# when commented out: File ended while scanning use of \textit . (ind and indname2 too)

```


```{r echo=FALSE, message= FALSE}
# define column headers for the pdf table
colnames = c("Domain", 
             "Indicator",
             "Time period", 
             "Latest value", "Time trend",
             "Latest value", "Time trend",
             "Latest value", "Time trend"
            )

```


```{r echo=FALSE, message= FALSE}

# get unique indicator list
unique_indicators <- unique(df$indname)

# Counter variable to keep track of which chart being generated
counter <- 1

# Iterate over each indicator and create/save a spineplot

for (indname in unique_indicators) {

  # Filter the data for the current indicator
  indicator_data <- df %>% filter(indname == !!indname)

  # create the plots
  maxval <- as.numeric(indicator_data$maxval)

  # All Scotland:
  # Unnest the trend column for plotting
  ind_df <- indicator_data %>%
    select(trend) %>%
    unnest(cols = c(trend))
  # create the plot
  if (length(ind_df$trend>0)) {
  spark1 <- create_sparkline_nat(ind_df, col1 = "#0078D4", col2 = "#B3D7F2", maxval)
  } else {
  spark1 <- plot_nodata()
  }
  # create filename
  filename1 <- sprintf("%s/scot_spark%s.png", graph_directory, counter)
  # Save the plot as a PNG with the filename created
  ggsave(filename1, plot = spark1, units="mm", width=15, height=5, dpi=300)

  # Most deprived quintile:
  # Unnest the trend column for plotting
  ind_df <- indicator_data %>%
    select(trend = mostdeptrend) %>%
    unnest(cols = c(trend))
  # create the plot
  if (length(ind_df$trend>0)) {
  spark2 <- create_sparkline_nat(ind_df, col1 = "#9B4393", col2 = "#E1C7DF",  maxval)
  } else {
  spark2 <- plot_nodata()
  }
  # create filename
  filename2 <- sprintf("%s/md_spark%s.png", graph_directory, counter)
  # Save the plot as a PNG with the filename created
  ggsave(filename2, plot = spark2, units="mm", width=15, height=5, dpi=300)

  # Least deprived quintile:
  # Unnest the trend column for plotting
  ind_df <- indicator_data %>%
    select(trend = leastdeptrend) %>%
    unnest(cols = c(trend))
   # create the plot
  if (length(ind_df$trend>0)) {
  spark3 <- create_sparkline_nat(ind_df, col1 = "#83BB26", col2 = "#DAEBBE", maxval)
  } else {
  spark3 <- plot_nodata()
  }
  # create filename
  filename3 <- sprintf("%s/ld_spark%s.png", graph_directory, counter)
  # Save the plot as a PNG with the filename created
  ggsave(filename3, plot = spark3, units="mm", width=15, height=5, dpi=300)

  # increase counter variable
  counter <- counter + 1
}



```



```{r, echo= FALSE, message = FALSE}

 # the final table to be displayed with charts embedded

table %>%
  group_by(domain) %>%
  knitr::kable(
    format = "latex", # must be in latex format for pdf download
    booktabs = TRUE, # generic formatting
    escape = FALSE, # ensures latex formatting on indicator column works (i.e.bold/italics,linebreaks)
    longtable = TRUE, # ensures table goes across multiple pages when too long
  #  align = c("l"), #left alignment within cells
    col.names = colnames
     ) %>% 
  # format header
  row_spec(0, background = "#032F4F", color = "white")  %>%
  add_header_above(., c(" " = 3,
                        "All Scotland" = 2,
                        "20% most deprived datazones" = 2,
                        "20% least deprived datazones" = 2),
                   escape = TRUE,
                   background = "#032F4F", color = "white") %>%
 # collapse_rows(columns = 1, valign = "top") %>%
  collapse_rows(columns = 1, valign = "top", row_group_label_position = "first", custom_latex_hline = 2:8, longtable_clean_cut = F) %>%
  # set column widths
  # column_spec(1, width = "8em") %>% # domain column
  # column_spec(2, width = "20em") %>% # indicator column
  # column_spec(3, width = "20em") %>% # value1
  # column_spec(4, width = "6em") %>% # trend1
  # column_spec(5, width = "3em") %>% # value2
  #  column_spec(6, width = "20em") %>% # trend2
  #  column_spec(7, width = "6em") %>% # value3
  #  column_spec(8, width = "3em") %>% # trend3
 # embed the chart images on each row of columns 5, 7 and 9
   kableExtra::column_spec(5, width = "6em",
                           image = sprintf("%s/scot_spark%s.png", graph_directory, table$id), color = "white") %>%
   kableExtra::column_spec(7, width = "6em",
                           image = sprintf("%s/md_spark%s.png", graph_directory, table$id), color = "white") %>%
   kableExtra::column_spec(9, width = "6em",
                           image = sprintf("%s/ld_spark%s.png", graph_directory, table$id), color = "white") %>%
  # repeats column headings when table goes across multiple pages # all give ! Extra alignment tab has been changed to \cr.
   kable_styling(latex_options = c("repeat_header")) %>%
  # add footnote with link to profiles tool
   footnote(general="\\\\url{https://scotland.shinyapps.io/phs-adult-mhi-dashboard/}",
            general_title = "Source: ", footnote_as_chunk = T, escape=F) %>%
   footnote(general="Alcohol-related hospitalisations due to mental and behavioural disorders only",
            general_title = "*", footnote_as_chunk = T, escape=F)
  # prevent domain from appearing on every row of column
 # kableExtra::collapse_rows(columns = 1, valign = "top", row_group_label_position = "first", custom_latex_hline = 2:8, longtable_clean_cut = F) 
# Warning: Error in [.data.frame: undefined columns selected



```


\elandscape